Making Sequential Programs Deterministic

### 2019-04-06 randoop-case-study.tex

To evaluate the usability of \theDeterminismChecker,
we applied it to the Randoop test
generator~\ifanonymous\citepalias{randoop-tool}\else\cite{PachecoLEB2007}\fi.
We also wrote specifications for libraries Randoop uses, such as the JDK, JUnit,
and others.
Like \theDeterminismCheckerImplementation,
all the case study materials are publicly available at \myurl{\todo{url for
    case study}} for reproducibility. \todo{This should include our
      our annotated plume-util, which we found was required to reproduce the
    results of the case study}

@@

We annotated one package at a time, starting with the most central packages
that are most depended upon.  Within a package, we followed a similar
strategy, annotating supertypes first.  If the determinism of classes and methods had been
documented, then our annotation effort
would have been easy, just converting English into type qualifiers.
Unfortunately, this was not the case: we had to reverse-engineer each
specification, largely from the methods it calls.  If a method $m$ calls a
method that we had previously annotated as returning a nondeterministic result, we annotated
$m$ as returning a nondeterministic result.
Sometimes after we
annotated a subclass, we had to go back and change annotations on
the supertype because the subclass implementation was
inconsistent with the specification we had guessed for the supertype.
After resolving most type-checking warnings by adding or changing
annotations, we manually examined each remaining one to determine whether
it was a bug or a false positive warning.
\todo{In some cases, the number of \<@Det> annotations was especially prevalent.
  In these cases we changed the default qualifier for that class to \<@Det>.}
The effort would have been much easier for someone familiar with Randoop,
and yet easier if done while code is being written and is malleable.

@@

\TheDeterminismChecker issued 37 false positive warnings, or 1 every 650
lines of code.

@@

The first two issues can be resolved by improvements to
\theDeterminismCheckerImplementation.
The third and fourth issues are limitations of the Java language, but a
local refactoring in the Randoop codebase can make the code type check.
Refactoring can also eliminate the last warning. \todo{Can it? I thought
  what happens when iterating over an @OrderNonDet collection is something we
can't verify with the checker.}

>>>>

To evaluate the usability of \theDeterminismChecker,
we applied it to the Randoop test
generator~\ifanonymous\citepalias{randoop-tool}\else\cite{PachecoLEB2007}\fi.
We also wrote specifications for libraries Randoop uses, such as the JDK,
JUnit, plume-util, and others.
Like \theDeterminismCheckerImplementation,
all the case study materials are publicly available at \myurl{\todo{url for
    case study}} for reproducibility.

@@

We annotated one package at a time, starting with the most central packages
that are most depended upon.  Within a package, we followed a similar
strategy, annotating supertypes first.  If the determinism of classes and methods had been
documented, then our annotation effort
would have been easy, just converting English into type qualifiers.
Unfortunately, this was not the case: we had to reverse-engineer each
specification, largely from the methods it calls.  If a method $m$ calls a
method that we had previously annotated as returning a nondeterministic result, we annotated
$m$ as returning a nondeterministic result.
Sometimes after we
annotated a subclass, we had to go back and change annotations on
the supertype because the subclass implementation was
inconsistent with the specification we had guessed for the supertype.
After resolving most type-checking warnings by adding or changing
annotations, we manually examined each remaining one to determine whether
it was a bug or a false positive warning.
When the number of \<@Det> annotations in a file was overwhelming,
we changed the default qualifier for that class to \<@Det>.  (Users
  can control defaulting on a file-by-file and method-by-method basis.)
The effort would have been much easier for someone familiar with Randoop,
and yet easier if done while code is being written and is malleable.

@@

\TheDeterminismChecker issued 37 false positive warnings, or 1 for every 650
lines of code.

@@

The first two issues can be resolved by improvements to
\theDeterminismCheckerImplementation.
The third and fourth issues are limitations of the Java language, but a
local refactoring in the Randoop codebase can make the code type check.
Refactoring can also eliminate the last warning, for example by using the
higher-order \<map()> function instead of a loop; alternately,
\theDeterminismChecker could be enhanced to pattern-match some loops.
